---
title: "Exercise12"
author: "Jessica V"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(tidyverse)
library(readr)
library(dtplyr)
library(MASS)
library(AICcmodavg)
library(MuMIn)
```



```{r}
f<- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Mammal_lifehistories_v2.txt"
d <- read_tsv(f, col_names = TRUE, show_col_types = FALSE) 
```
Do a bit of exploratory data analysis with this dataset, e.g., using the {skimr} package. 
Which of the variables are categorical and which are numeric?

Numerica columns are : "mass(g)"  "gestation(mo)" "newborn(g)"    "weaning(mo)"   "wean mass(g)"  "AFR(mo)"       "max. life(mo)" "litter size"   "litters/year" and "refs".

Character  columns are : "order"   "family"  "Genus" and  "species".
```{r}

head(d)

summary(d)

#Here I view the classes of all of the columns
sapply(d, class)

#Here I extract only numeric columns
numericCols<-names(d)[sapply(d, is.numeric)]
 
numericCols
#Here I extract only character columns
categoricalCols<-names(d)[sapply(d, is.character)]

categoricalCols
```

# Challenge

#Step 1

Replace all values of -999 (the authors’ code for missing data) with NA.
```{r}
d[d== -999] <- NA
head(d)
```

#Step 2

Drop the variables "litter size" and "refs".

```{r}
colnames(d)

newd <- d[, -c(12,14)]

colnames(newd)
```

#Step 3

Log transform all of the other numeric variables.

```{r}

lognewd <- newd %>%
 mutate(across(where(is.numeric), ~log(.+1)))

head(lognewd)
```
#Step 4

Regress the age [gestation(mo), weaning(mo), AFR(mo), and max. life(mo)] and mass [newborn(g) and wean mass(g)] variables on overall body mass(g) and add the residuals to the dataframe as new variables (relGest, relWean, relAFR, relLife, relNewbornMass, and relWeaningMass).

```{r}
colnames(newd)

age <- lm(cbind(`gestation(mo)`, `weaning(mo)`, `AFR(mo)`, `max. life(mo)`) ~ `mass(g)`, data=newd, na.action=na.exclude)

mass <- lm(cbind(`newborn(g)`, `wean mass(g)`) ~ `mass(g)`, data =newd, na.action=na.exclude)

newd$relGest <- residuals(age)[, "gestation(mo)"]

newd$relWean <- residuals(age)[, "weaning(mo)"]

newd$relAFR <- residuals(age)[, "AFR(mo)"]

newd$relLife <- residuals(age)[,"max. life(mo)"]

newd$relNewbornMass <- residuals(mass)[,"newborn(g)"]
  
newd$relWeaningMass <- residuals(mass)[,"wean mass(g)"]

head(newd)
colnames(newd)
```

#Step 5
 
Plot residuals of max lifespan (relLife) in relation to Order. Which mammalian orders have the highest residual lifespan?
Plot P:
Probiscidea has the highest avg residual lifespan, followed by Cetacea and Sirenia.

Plot residuals of newborn mass (relNewbornMass) in relation to Order. Which mammalian orders have the have highest residual newborn mass?
Plot P1:
Proboscidea has thehigest residual newborn mass, followed by perissodactyla.

Plot residuals of weaning mass (relWeaningMass) in relation to Order. Which mammalian orders have the have highest relative weaning mass?
Plot P2:
Perissodactyla and Proboscidae


```{r}

p <- ggplot(data = newd, aes(x = `order`, y = `relLife` )) + geom_boxplot() + xlab("Order") +
    theme(axis.text.x = element_text(angle = 90))


plot(p)

p1 <- ggplot(data = newd, aes(x = `order`, y = `relNewbornMass` )) + geom_boxplot() + xlab("Order") +
    theme(axis.text.x = element_text(angle = 90))

plot(p1)

p2 <- ggplot(data = newd, aes(x = `order`, y = `relWeaningMass` )) + geom_boxplot() + xlab("Order") +
    theme(axis.text.x = element_text(angle = 90))

plot(p2)
```

#Step 6

Run models and a model selection process to evaluate what (now log transformed) variables best predict each of the two response variables, max. life(mo) and AFR(mo), from the set of the following predictors: gestation(mo), newborn(g), weaning(mo), wean mass(g), litters/year, and overall body mass(g).



For each of the two response variables, indicate what is the best model overall based on AICc and how many models have a delta AICc of 4 or less?

What variables, if any, appear in all of this set of “top” models?
Calculate and plot the model-averaged coefficients and their CIs across this set of top models.
```{r}

drop_na(newd)
colnames(newd)
#FOR max. life (mo)
m <- lm(data = newd, `max. life(mo)` ~ `gestation(mo)`+ `newborn(g)`+ `weaning(mo)`+ `wean mass(g)`+ `litters/year`, + `mass(g)`) 

(s <- stepAIC(m, scope = . ~ ., direction = "both"))
summary (s)

#For AFR(mo)
m2 <- lm(data = newd, `AFR(mo)` ~ `gestation(mo)`+ `newborn(g)`+ `weaning(mo)`+ `wean mass(g)`+ `litters/year`, + `mass(g)`) 

(s2 <- stepAIC(m2, scope = . ~ ., direction = "both"))

summary(s2)
```

```{r}
detach(package:MASS)
library(AICcmodavg)
#aictab(list(m0, m1, m2), c("m0", "m1", "m2"))
detach(package:AICcmodavg)


#(mods.avg <- summary(model.avg(mods, subset = delta < 4, fit = TRUE)))


```
